---
title: "Transition Aged Youth Inflow"
author: "Genelle Denzin"
date: "6/3/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(lubridate)
library(here)

reportyear <- 2019
# BEFORE KNITTING THIS, be sure you're running 00_get_the_CSV_things.R with the
# correct directory

load(here("images/YOHMIS.RData"))
```



```{r tay_inflow, include=FALSE}
 
 inflow <- Enrollment %>%
   filter(RelationshipToHoH == 1 &
            entered_between(., paste("0101", reportyear),
                            paste("1231", reportyear))) %>%
   select(HouseholdID) %>%
   left_join(Enrollment,
             by = c("HouseholdID")) %>%
   group_by(HouseholdID) %>%
   summarise(oldest = max(AgeAtEntry)) %>%
   ungroup() %>%
   filter(oldest > 17 & oldest < 25) %>%
   left_join(Enrollment, by = c("HouseholdID")) %>%
   select(
     HouseholdID,
     PersonalID,
     RelationshipToHoH,
     EntryDate,
     ExitDate,
     ProjectName,
     AgeAtEntry
   )
 deduplicated_clients <- inflow %>%
   select(PersonalID) %>%
   unique() %>%
   count() %>%
   pull()
```

## Total Transition Aged Youth Inflow January `r reportyear` to December `r reportyear`

Transition Aged Youth includes households where the oldest person in the household is between the age of 18 and 24. **`r deduplicated_clients`** total unique clients in transition aged youth households entered a project in the Mahoning CoC during `r reportyear`.

## Household Types

A household is a unit of people presenting for services to be housed together. Here we will refer to a household with more than one person in it as Families (regardless of relations) and households with a single person in it as Individuals. 

```{r hh_types}

family_units <- inflow %>%
   filter(str_detect(HouseholdID, fixed("h_"))) %>%
   select(HouseholdID) %>%
   unique() %>%
   count() %>%
   pull()

clients_in_families <- inflow %>%
   filter(str_detect(HouseholdID, fixed("h_"))) %>%
   select(PersonalID) %>%
   unique() %>% 
   count() %>%
   pull()

individuals <- inflow %>%
   filter(str_detect(HouseholdID, fixed("s_"))) %>%
   select(PersonalID) %>%
   unique() %>%
   count() %>%
   pull()

```

During `r reportyear`: 

- `r clients_in_families` clients in `r family_units` Families headed by a transition aged youth entered a project in the Mahoning County CoC.
- `r individuals` transition aged youth Individuals entered a project in the Mahoning County CoC.

*Some clients may have presented for housing as an Individual and, in another episode, as part of a Family. Those clients are being counted in both household types.*

## Fleeing Domestic Violence

```{r}

dv <- inflow %>%
   left_join(
      HealthAndDV %>%
         filter(DataCollectionStage == 1) %>%
         select(PersonalID,
                CurrentlyFleeing,
                WhenOccurred),
      by = c("PersonalID")
   ) %>%
   mutate(
      CurrentlyFleeing = if_else(is.na(CurrentlyFleeing), 99, CurrentlyFleeing),
      WhenOccurred = if_else(is.na(WhenOccurred), 99, WhenOccurred),
      CurrentlyFleeing = case_when(
         CurrentlyFleeing %in% c(0, 99) &
            WhenOccurred %in% c(4, 8, 9, 99) ~ "No",
         CurrentlyFleeing == 1 |
            WhenOccurred %in% c(1:3) ~ "Yes",
         CurrentlyFleeing %in% c(8, 9, 99) |
            WhenOccurred %in% c(8, 9, 99) ~ "Unknown"
      )
   ) %>%
   select(HouseholdID,
          PersonalID,
          EntryDate,
          RelationshipToHoH,
          AgeAtEntry,
          CurrentlyFleeing)

households_dv <- dv %>%
   filter(CurrentlyFleeing == "Yes") %>%
   select(HouseholdID) %>%
   unique() %>%
   count() %>%
   pull()

```

During `r reportyear`, `r households_dv` households of one or more who were fleeing domestic violence and headed by a transition aged youth entered a project in the Mahoning County CoC.



