---
title: "Children 2017 - 2018"
author: "Genelle Denzin"
date: "March 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

### Served in ES, TH, or SH During the Year Excluding Unsheltered Homeless

```{r basicdatasets}
library(lubridate)
library(dplyr)
library(gt)
library(janitor)
load("data/COHHIOHMIS.RData")
smallEnrollment <- Enrollment %>% 
  select(EnrollmentID, PersonalID, EntryDate, ProjectID)
smallExit <- Exit %>%
  select(EnrollmentID, ExitDate)
smallProject <- Project %>%
  select(ProjectID, ProjectType)
smallClient <- Client %>%
  select(PersonalID, DOB)
rm(Affiliation, Disabilities, EmploymentEducation, EnrollmentCoC, Export, Funder, Geography, HealthAndDV, IncomeBenefits, Inventory, Organization, ProjectCoC, Scores, Services, Users)
```



```{r ESTHSH}
ESTHSH <- left_join(smallEnrollment, smallClient, by = "PersonalID") %>%
  left_join(., smallExit, by = "EnrollmentID") %>%
  left_join(., smallProject, by = "ProjectID") %>%
  mutate(AgeIn2017 = age_years(DOB, ymd("2017-01-01")),
         AgeIn2018 = age_years(DOB, ymd("2018-01-01")),
         ServedDuring2017 = ymd(EntryDate) < ymd("2018/01/01") &
           (is.na(ExitDate) | ymd(ExitDate) >= ymd("2017/01/01")),
         ServedDuring2018 = ymd(EntryDate) < ymd("2019/01/01") &
           (is.na(ExitDate) | ymd(ExitDate) >= ymd("2018/01/01"))) %>% 
  filter(((AgeIn2017 >=0 & AgeIn2017 < 18) | 
            (AgeIn2018 >=0 & AgeIn2018 < 18)) &
                 ProjectType %in% c(1, 2, 8) &
                 (ServedDuring2017 == TRUE | ServedDuring2018 == TRUE)) %>%
  select(PersonalID, AgeIn2017, AgeIn2018, ServedDuring2017, ServedDuring2018)

Year2017ESTHSH <- ESTHSH %>%
  filter(AgeIn2017 < 18 & AgeIn2017 >= 0 & ServedDuring2017 == TRUE) %>%
  select(-AgeIn2018, -ServedDuring2018) %>%
  mutate(Infant = AgeIn2017 < 1) %>%
  group_by(Infant) %>%
  summarise(ServedIn2017 = n()) %>%
  ungroup()
Year2018ESTHSH <- ESTHSH %>%
  filter(AgeIn2018 < 18 & AgeIn2018 >= 0 & ServedDuring2018 == TRUE) %>%
  select(-AgeIn2017, -ServedDuring2017) %>%
  mutate(Infant = AgeIn2018 < 1) %>%
  group_by(Infant) %>%
  summarise(ServedIn2018 = n()) %>%
  ungroup()

gt(Year2017ESTHSH)
gt(Year2018ESTHSH)
```



### Served in ES, TH, SH, or Outreach During the Year Including Unsheltered Homeless

```{r ESTHSHOUT}
ESTHSHOUT <- left_join(smallEnrollment, smallClient, by = "PersonalID") %>%
  left_join(., smallExit, by = "EnrollmentID") %>%
  left_join(., smallProject, by = "ProjectID") %>%
  mutate(AgeIn2017 = age_years(DOB, ymd("2017-01-01")),
         AgeIn2018 = age_years(DOB, ymd("2018-01-01")),
         ServedDuring2017 = ymd(EntryDate) < ymd("2018/01/01") &
           (is.na(ExitDate) | ymd(ExitDate) >= ymd("2017/01/01")),
         ServedDuring2018 = ymd(EntryDate) < ymd("2019/01/01") &
           (is.na(ExitDate) | ymd(ExitDate) >= ymd("2018/01/01"))) %>% 
  filter(((AgeIn2017 >=0 & AgeIn2017 < 18) | 
            (AgeIn2018 >=0 & AgeIn2018 < 18)) &
                 ProjectType %in% c(1, 2, 8, 4) &
                 (ServedDuring2017 == TRUE | ServedDuring2018 == TRUE)) %>%
  select(PersonalID, AgeIn2017, AgeIn2018, ServedDuring2017, ServedDuring2018)

Year2017ESTHSHOUT <- ESTHSHOUT %>%
  filter(AgeIn2017 < 18 & AgeIn2017 >= 0 & ServedDuring2017 == TRUE) %>%
  select(-AgeIn2018, -ServedDuring2018) %>%
  mutate(Infant = AgeIn2017 < 1) %>%
  group_by(Infant) %>%
  summarise(ServedIn2017 = n()) %>%
  ungroup()
Year2018ESTHSHOUT <- ESTHSHOUT %>%
  filter(AgeIn2018 < 18 & AgeIn2018 >= 0 & ServedDuring2018 == TRUE) %>%
  select(-AgeIn2017, -ServedDuring2017) %>%
  mutate(Infant = AgeIn2018 < 1) %>%
  group_by(Infant) %>%
  summarise(ServedIn2018 = n()) %>%
  ungroup()

gt(Year2017ESTHSHOUT)
gt(Year2018ESTHSHOUT)
```

I could not get this to work with the Infants being a subset of all the children. I need to learn how to add a total row in dplyr, how can this be that hard? Anyway, I just pulled the numbers out into Excel on the HMIS Share under Reporting > Random Report Requests > Children 2017 2018.