---
title: "LoTH Unsheltered"
author: "Genelle Denzin"
date: "5/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
```

## Data to Help with Extrapolating Length of Time Homeless data for Unsheltered

This is looking at all Unsheltered adults with Entries in 2019 who were entered by anyone on the CoC team. The following is a Summary of the table we are using for the following analysis.

```{r summary}

load("data/COHHIOHMIS.RData")

unshelteredLoTH <- Enrollment %>%
  select(EnrollmentID, 
         PersonalID, 
         EntryDate, 
         ProjectID,
         AgeAtEntry,
         UserCreating,
         DateToStreetESSH, 
         TimesHomelessPastThreeYears,
         MonthsHomelessPastThreeYears) %>%
  filter(ProjectID == 1695 &
           AgeAtEntry > 17,
           UserCreating %in% c("Mulryan, Erica/@COHHIO(1239)",
                               "Basting, Hannah/@COHHIO(1563)",
                               "Sechang, Sandy/@COHHIO(1426)",
                               "Walton, Valerie/@COHHIO(1628)",
                               "Hoffman, Carolyn/@COHHIO(1624)") &
           ymd(EntryDate) > mdy(01012019))
summary(unshelteredLoTH)
```

## Times Homeless 

Distribution of the Times Homeless data element. 1 = 1 time, 2, = 2 times, 3 = 3 times, and 4 = 4 or more times. I excluded the missings and DKRs.

```{r timesplot}
ggplot(filter(unshelteredLoTH, !TimesHomelessPastThreeYears %in% c(99, 8, 9) &
                !is.na(TimesHomelessPastThreeYears)), aes(TimesHomelessPastThreeYears)) +
  geom_histogram(binwidth = 1)
```

## Months Homeless 

Distribution of the Months Homeless data element. 101 = 1 month, 102, = 2 months, ... 113 = more than 12 months. I excluded the missings and DKRs.

```{r monthsplot}
ggplot(filter(unshelteredLoTH, !MonthsHomelessPastThreeYears %in% c(8, 9, 99) &
                !is.na(MonthsHomelessPastThreeYears)), aes(MonthsHomelessPastThreeYears)) +
  geom_histogram(binwidth = 1)
```


