---
title: "Scores by Geography"
author: "Genelle Denzin"
date: "March 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(janitor)
library(gt)
load("data/COHHIOHMIS.RData")
rm(Affiliation, EmploymentEducation, Export, Funder, HealthAndDV, IncomeBenefits,
   Geography, Inventory, Organization, ProjectCoC, Services)
```

## Average VI-SPDAT Scores at the Provider Level

The following table counts each Entry into a PH project in a County during the reporting period for Heads of Household only. It is very important that your Duplicate Entry Exits and your Household Data Quality tabs are totally clear for this report to be accurate. It is also important that your VI-SPDAT scores are ON THE HEAD OF HOUSEHOLD'S RECORD. Any scores recorded on non-HoHs will not be counted here. Any Heads of Household who entered your RRH or PSH project without a score will be counted as having a score of 0. Also if a HoH is missing their County data or they were served in a County outside the Ohio Balance of State, they will also not show here.

```{r scores}
smallEnrollment <- Enrollment %>% 
  left_join(Project, by = "ProjectID") %>%
  select(EnrollmentID, 
         PersonalID,
         ProjectID,
         ProjectType,
         ProjectName,
         OperatingStartDate,
         OperatingEndDate,
         EntryDate,
         HouseholdID,
         RelationshipToHoH,
         MoveInDate,
         CountyServed)
Entries <- smallEnrollment %>%
  filter(
  ProjectType %in% c(3, 9, 13) &
  ymd(OperatingStartDate) < ymd("2019/04/01") &
    (ymd(OperatingEndDate) > ymd("2019/01/01") |
       is.na(OperatingEndDate)) &
  ymd(EntryDate) > ymd("2019/01/01") &
  ymd(EntryDate) < ymd("2019/04/01")
  )
SPDATsByProject <- left_join(Entries, Scores, by = "PersonalID") %>%
  select(-ProjectType,
         -OperatingStartDate,
         -OperatingEndDate,
         -SPDATRecordID,
         -SPDATProvider) %>%
  filter(
    RelationshipToHoH == 1 &
    (ymd(StartDate) <= ymd(EntryDate) | is.na(StartDate)) &
      !CountyServed %in% c(
        "Montgomery",
        "Cuyahoga",
        "Mahoning",
        "Lucas",
        "Stark",
        "Summit",
        "Hamilton",
        "Franklin",
        "--Outside of Ohio--"
        ) &
    !is.na(CountyServed)
    ) %>%
  group_by(EnrollmentID) %>%
  mutate(MaxScoreDate = max(ymd(StartDate))) %>%
  filter(ymd(StartDate) == ymd(MaxScoreDate) | is.na(StartDate)) %>%
  mutate(MaxScore = max(Score)) %>%
  filter(Score == MaxScore | is.na(StartDate)) %>%
  distinct() %>%
  ungroup() %>%
  select(-MaxScoreDate, -MaxScore) %>%
  mutate(ScoreAdjusted = if_else(is.na(Score), 0, Score)) %>%
  filter(!is.na(ScoreAdjusted))
# I think this should be based on EnrollmentID not ClientID because each decision
# to enroll a household should rely on a SPDAT score
ProviderAverages <- SPDATsByProject %>%
  select(EnrollmentID, ProjectName, ScoreAdjusted) %>%
  group_by(ProjectName) %>%
  summarise(AverageScore = round(mean(ScoreAdjusted), 1),
            EnrollmentCount = n())

CountyHousedAverageScores <- SPDATsByProject %>%
  group_by(CountyServed) %>%
  summarise(HousedAverageScore = round(mean(ScoreAdjusted),1),
            HHsHousedInCounty = n())

gt(CountyHousedAverageScores)

```

## Duplicate Enrollments Affecting This Report

If you see your Project listed here, please run your Data Quality 1 report and correct your Duplicate Entry Exits tab.
```{r}
duplicate_enrollments <- get_dupes(SPDATsByProject, PersonalID, ProjectID)

gt(duplicate_enrollments)
```

## VI-SPDAT Scores Saved to Non-Heads-of-Households

If you have clients here, you should either verify the scores saved here are valid or the correct client is marked as the Head of Household.

```{r}
SPDATsOnNonHoHs <- left_join(Entries, Scores, by = "PersonalID") %>%
  filter(RelationshipToHoH != 1 & !is.na(Score)) %>%
  select(ProjectName, PersonalID, EntryDate, Score) %>%
  arrange(ProjectName)
gt(SPDATsOnNonHoHs)
```

## County Averages

The Heads of Households counted here were served and scored in a ES, TH, SH, or Outreach project during the reporting period and the County they were served in was in the Ohio Balance of State. If a Head of Household had a County outside the Balance of State or if that data was missing, they are not being counted.

```{r}
SPDATsByCounty <-
  left_join(smallEnrollment, Scores, by = "PersonalID") %>%
  filter(
    ProjectType %in% c(1, 2, 4, 8) &
    RelationshipToHoH == 1 & 
    ymd(OperatingStartDate) < ymd("2019/04/01") &
     (ymd(OperatingEndDate) > ymd("2019/01/01") 
      | is.na(OperatingEndDate)) &
    ymd(EntryDate) > ymd("2019/01/01") & 
    ymd(EntryDate) < ymd("2019/04/01") & 
    (ymd(StartDate) <= ymd(EntryDate)) & 
    !CountyServed %in% c("Montgomery", 
                         "Cuyahoga", 
                         "Mahoning", 
                         "Lucas", 
                         "Stark", 
                         "Summit", 
                         "Hamilton", 
                         "Franklin",
                         "--Outside of Ohio--") &
      !is.na(CountyServed)) %>%
  group_by(PersonalID) %>%
  mutate(MaxEntry = max(ymd(EntryDate))) %>% # most recent EE
  filter(ymd(MaxEntry) == ymd(EntryDate)) %>%  
  mutate(MaxScoreDate = max(ymd(StartDate))) %>% # most recent score
  filter(ymd(StartDate) == ymd(MaxScoreDate)) %>%
  mutate(MaxScore = max(Score)) %>% # highest score
  filter(Score == MaxScore) %>%
  select(CountyServed, PersonalID, Score) %>%
  distinct() %>%
  ungroup()

CountyAverageScores <- SPDATsByCounty %>%
  group_by(CountyServed) %>%
  summarise(AverageScore = round(mean(Score), 1), 
            HHsLHinCounty = n())

gt(CountyAverageScores)
```

## Compare Average Scores Literally Homeless to Housed in a County

The following compares the Average VI-SPDAT Score of those who have entered a project that would mean they are literally homeless within a County to the Average VI-SPDAT Score of those who have an Entry into a Permanent Housing project type (RRH or PSH).

```{r}
Regions <- read_csv("data/Regions.csv") %>%
  mutate(RegionName = paste0("Homeless Planning Region ", Region))

Compare <- 
  full_join(CountyAverageScores, 
            CountyHousedAverageScores, 
            by = "CountyServed") %>%
  arrange(CountyServed) %>%
  left_join(., Regions, by = c("CountyServed" = "County")) %>%
  group_by(RegionName)

ggplot(Compare %>% filter(Region == 15), 
       aes(x = CountyServed, y = AverageScore)) + 
  geom_point(aes(x = CountyServed, y = AverageScore), size = 10, shape = 95) +
  theme(axis.text.x = element_text(size = 10)) +
  geom_point(aes(x = CountyServed, y = HousedAverageScore), 
             size = 4, 
             shape = 17) +
  xlab(Compare$RegionName)
```

