---
title: "LH DV Served in 2018"
author: "Genelle Denzin"
date: "7/17/2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(lubridate)
library(here)
load(here("images/COHHIOHMIS.RData"))
```

## Literally Homeless Domestic Violence Survivors Served in 2018

The following represents clients who indicated they were "Currently Fleeing" at Entry into any Ohio Balance of State project that enters into HMIS.


```{r summary}

dvClients <- HealthAndDV %>%
  select(PersonalID, EnrollmentID, DomesticViolenceVictim, WhenOccurred,
         CurrentlyFleeing) %>%
  filter(CurrentlyFleeing == 1) %>%
  left_join(Enrollment, by = c("PersonalID", "EnrollmentID")) %>% 
  select(EnrollmentID, PersonalID, ProjectType, AgeAtEntry, Destination, CountyServed, EntryDate, ExitDate) %>%
  filter(served_between(., "01012018", "12312018")) %>%
  select(-EntryDate, -ExitDate)

nrow(dvClients)
n_distinct(dvClients$PersonalID)

distinctClients <- HealthAndDV %>%
  filter(CurrentlyFleeing == 1) %>%
  group_by(PersonalID) %>%
  summarise(Enrollments = n()) %>%
  ungroup()

repeaters <- distinctClients %>%
  filter(Enrollments > 1) %>%
  summarise(AvgNoEnrollments = mean(Enrollments))

repeaters

```

## By Project Type

More clients who are currently fleeing domestic violence present at Emergency Shelters than Rapid Rehousing, but

```{r project_type}

dvProjectTypes <- dvClients %>%
  mutate(ProjectType = case_when(
    ProjectType == 1 ~ "Emergency Shelter",
    ProjectType == 2 ~ "Transitional Housing",
    ProjectType == 3 ~ "Permanent Supportive Housing",
    ProjectType == 4 ~ "Street Outreach",
    ProjectType == 6 ~ "Services Only",
    ProjectType == 7 ~ "Other",
    ProjectType == 8 ~ "Safe Haven",
    ProjectType == 9 ~ "PSH- Housing Only",
    ProjectType == 12 ~ "Prevention",
    ProjectType == 13 ~ "Rapid Rehousing",
    ProjectType == 14 ~ "Other"
  )) %>%
  group_by(ProjectType) %>%
  summarise(DVSurvivors = n()) %>%
  ungroup() %>%
  arrange(desc(DVSurvivors))

dvProjectTypes
  
```

# Where are these clients going after they exit?

```{r destinations}
dvDestinations <-  dvClients %>%
  mutate(DestinationGroup = case_when(
    Destination %in% c(3, 10:11, 19:23, 26, 28, 31) ~ "Permanent",
    Destination %in% c(1:2, 12:14, 16, 18, 27) ~ "Temporary",
    Destination %in% c(4:7, 15, 25) ~ "Institutional",
    Destination %in% c(8:9, 17, 24, 29) ~ "Other",
    Destination %in% c(30, 99) ~ "Not Collected",
    is.na(Destination) ~ "Still in Project"
  )) %>%
  group_by(DestinationGroup) %>%
  summarise(DVSurvivors = n()) %>%
  ungroup() %>%
  arrange(desc(DVSurvivors))

dvDestinations

```
  
## Geography of DV Entries into the Homeless System
  
```{r geo}

dvClients %>%
  group_by(CountyServed) %>%
  summarise(DVSurvivors = n()) %>%
  ungroup() %>%
  arrange(desc(DVSurvivors))

```
  
## Age Demographics

```{r age}
dvClients %>% 
  group_by(PersonalID) %>%
  summarise(FirstEntryAge = min(AgeAtEntry)) %>%
  ungroup() %>%
  mutate(AgeBucket = case_when(
    FirstEntryAge < 6 ~ "0-5",
    between(FirstEntryAge, 6, 18) ~ "6-18",
    between(FirstEntryAge, 19, 25) ~ "19-25",
    between(FirstEntryAge, 26, 50) ~ "26-50",
    between(FirstEntryAge, 51, 70) ~ "51-70",
    FirstEntryAge > 70 ~ "Over 70"
  ),
  AgeBucket = fct_reorder(AgeBucket, FirstEntryAge)) %>%
  group_by(AgeBucket) %>%
  summarise(DVSurvivors = n())
  
```



