---
title: "LH DV Served in 2018"
author: "Genelle Denzin"
date: "7/17/2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(lubridate)
library(here)
load(here("images/COHHIOHMIS.RData"))
```

## Literally Homeless Domestic Violence Survivors Served in 2018

The following represents clients who indicated they were "Currently Fleeing" at Entry into any Ohio Balance of State project that enters into HMIS.


```{r summary}

dvClients <- HealthAndDV %>%
  select(PersonalID, EnrollmentID, DomesticViolenceVictim, WhenOccurred,
         CurrentlyFleeing) %>%
  filter(CurrentlyFleeing == 1) %>%
  left_join(Enrollment, by = c("PersonalID", "EnrollmentID")) %>% 
  select(EnrollmentID, PersonalID, ProjectType, AgeAtEntry, Destination, CountyServed, EntryDate, ExitDate) %>%
  filter(served_between(., "01012018", "12312018")) %>%
  select(-EntryDate, -ExitDate)

distinctClientsESRRHTH <- dvClients %>%
  filter(ProjectType %in% c(1:2, 13)) 

counts <- c(nrow(dvClients), n_distinct(dvClients$PersonalID), n_distinct(distinctClientsESRRHTH$PersonalID))
names <- c("Entries", "Distinct Clients", "Distinct Clients in ES, TH, RRH")

cbind(names, counts)

distinctClients <- dvClients %>%
  group_by(PersonalID) %>%
  summarise(Enrollments = n()) %>%
  ungroup()

repeaters <- distinctClients %>%
  filter(Enrollments > 1) %>%
  summarise(AvgNoEnrollments = mean(Enrollments))

repeaters

```
When clients who are fleeing domestic violence come into our system, the ones who had more than one project stay during the year entered a project an average of 3 times. This would include entering and re-entering the same project.

## By Project Type

A client can show in multiple categories here.

```{r project_type}

dvProjectTypes <- dvClients %>%
  mutate(ProjectType = case_when(
    ProjectType == 1 ~ "Emergency Shelter",
    ProjectType == 2 ~ "Transitional Housing",
    ProjectType == 3 ~ "Permanent Supportive Housing",
    ProjectType == 4 ~ "Street Outreach",
    ProjectType == 6 ~ "Services Only",
    ProjectType == 7 ~ "Other",
    ProjectType == 8 ~ "Safe Haven",
    ProjectType == 9 ~ "PSH- Housing Only",
    ProjectType == 12 ~ "Prevention",
    ProjectType == 13 ~ "Rapid Rehousing",
    ProjectType == 14 ~ "Other"
  )) %>%
  group_by(ProjectType) %>%
  summarise(Fleeing = n()) %>%
  ungroup() %>%
  arrange(desc(Fleeing))

dvProjectTypes
  
```

# Where are these clients going after they exit?

A client can show up in multiple categories here.

```{r destinations}
dvDestinations <-  dvClients %>%
  mutate(DestinationGroup = case_when(
    Destination %in% c(3, 10:11, 19:23, 26, 28, 31) ~ "Permanent",
    Destination %in% c(1:2, 12:14, 16, 18, 27) ~ "Temporary",
    Destination %in% c(4:7, 15, 25) ~ "Institutional",
    Destination %in% c(8:9, 17, 24, 29) ~ "Other",
    Destination %in% c(30, 99) ~ "Not Collected",
    is.na(Destination) ~ "Still in Project"
  )) %>%
  group_by(DestinationGroup) %>%
  summarise(Fleeing = n()) %>%
  ungroup() %>%
  arrange(desc(Fleeing))

dvDestinations

```
  
## Geography of DV Entries into the Homeless System

A client can show up in multiple categories here.
  
```{r geo}

dvClients %>%
  group_by(CountyServed) %>%
  summarise(Fleeing = n()) %>%
  ungroup() %>%
  arrange(desc(Fleeing))

```
  
## Age Demographics

Unique clients.

```{r age}
dvClients %>% 
  group_by(PersonalID) %>%
  summarise(FirstEntryAge = min(AgeAtEntry)) %>%
  ungroup() %>%
  mutate(AgeBucket = case_when(
    FirstEntryAge < 6 ~ "0-5",
    between(FirstEntryAge, 6, 18) ~ "6-18",
    between(FirstEntryAge, 19, 25) ~ "19-25",
    between(FirstEntryAge, 26, 50) ~ "26-50",
    between(FirstEntryAge, 51, 70) ~ "51-70",
    FirstEntryAge > 70 ~ "Over 70"
  ),
  AgeBucket = fct_reorder(AgeBucket, FirstEntryAge)) %>%
  group_by(AgeBucket) %>%
  summarise(Fleeing = n())
  
```



