---
title: "Scores by Geography"
author: "Genelle Denzin"
date: "March 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
#THIS IS NOT A FINAL DOCUMENT. Please see the logic in R minor and 02_QPR_SPDATs.R in this project for updated logic.
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(janitor)
library(gt)
load("images/COHHIOHMIS.RData")
rm(Affiliation, EmploymentEducation, Export, Funder, HealthAndDV, Disabilities, IncomeBenefits, Geography, Inventory, Offers, Organization, ProjectCoC, Services, VeteranCE)
```

## County Averages

The Heads of Households counted here were served and scored in a ES, TH, SH, or Outreach project during the reporting period and the County they were served in was in the Ohio Balance of State. If a Head of Household had a County outside the Balance of State or if that data was missing, they are not being counted. When there are multiple entries for the same client, this only counts the most recent entry. When there are multiple scores, this only counts the most recent score. There should not be more than 1 score on the same day, but if there are it is counting the highest score.

```{r countyscores}
ReportStart <- "01012019"
ReportEnd <- "03312019"

smallEnrollment <- Enrollment %>% 
  left_join(Project, by = "ProjectID") %>%
  select(EnrollmentID, 
         PersonalID,
         ProjectID,
         ProjectType,
         ProjectName,
         OperatingStartDate,
         OperatingEndDate,
         EntryDate,
         ExitDate,
         RelationshipToHoH,
         CountyServed)

Entries <- smallEnrollment %>%
  filter(
  ProjectType %in% c(3, 9, 13) &
  operating_between(smallEnrollment, ReportStart, ReportEnd) &
  served_between(smallEnrollment, ReportStart, ReportEnd)
  )

SPDATsByCounty <-
  left_join(smallEnrollment, Scores, by = "PersonalID") %>%
  filter(
    ProjectType %in% c(1, 2, 4, 8) &
    RelationshipToHoH == 1 & 
    ymd(StartDate) <= ymd(EntryDate) & 
    !CountyServed %in% c("Montgomery", 
                         "Cuyahoga", 
                         "Mahoning", 
                         "Lucas", 
                         "Stark", 
                         "Summit", 
                         "Hamilton", 
                         "Franklin",
                         "--Outside of Ohio--") &
      !is.na(CountyServed)) %>%
  group_by(PersonalID) %>%
  mutate(MaxEntry = max(ymd(EntryDate))) %>% # most recent EE
  filter(ymd(MaxEntry) == ymd(EntryDate)) %>%  
  mutate(MaxScoreDate = max(ymd(StartDate))) %>% # most recent score
  filter(ymd(StartDate) == ymd(MaxScoreDate)) %>%
  mutate(MaxScore = max(Score)) %>% # highest score
  filter(Score == MaxScore) %>%
  select(CountyServed, PersonalID, Score) %>%
  distinct() %>%
  ungroup()

CountyAverageScores <- SPDATsByCounty %>%
  group_by(CountyServed) %>%
  summarise(AverageScore = round(mean(Score), 1), 
            HHsLHinCounty = n())

gt(CountyAverageScores)
```

## Average VI-SPDAT Scores at the Provider Level

The following table counts each Entry into a PH project in a County during the reporting period for Heads of Household only. It is very important that your Duplicate Entry Exits and your Household Data Quality tabs are totally clear for this report to be accurate. It is also important that your VI-SPDAT scores are ON THE HEAD OF HOUSEHOLD'S RECORD. Any scores recorded on non-HoHs will not be counted here. Any Heads of Household who entered your RRH or PSH project without a score will be counted as having a score of 0. Also if a HoH is missing their County data or they were served in a County outside the Ohio Balance of State, they will also not show here.

```{r projectscores}

SPDATsByProject <- left_join(Entries, Scores, by = "PersonalID") %>%
  select(-ProjectType,
         -OperatingStartDate,
         -OperatingEndDate,
         -SPDATRecordID,
         -SPDATProvider) %>%
  filter(
    RelationshipToHoH == 1 &
    (ymd(StartDate) <= ymd(EntryDate) | is.na(StartDate)) &
      !CountyServed %in% c(
        "Montgomery",
        "Cuyahoga",
        "Mahoning",
        "Lucas",
        "Stark",
        "Summit",
        "Hamilton",
        "Franklin",
        "--Outside of Ohio--"
        ) &
    !is.na(CountyServed)
    ) %>%
  group_by(EnrollmentID) %>%
  mutate(MaxScoreDate = max(ymd(StartDate))) %>%
  filter(ymd(StartDate) == ymd(MaxScoreDate) | is.na(StartDate)) %>%
  mutate(MaxScore = max(Score)) %>%
  filter(Score == MaxScore | is.na(StartDate)) %>%
  distinct() %>%
  ungroup() %>%
  select(-MaxScoreDate, -MaxScore) %>%
  mutate(ScoreAdjusted = if_else(is.na(Score), 0, Score)) %>%
  filter(!is.na(ScoreAdjusted))

ProviderAverages <- SPDATsByProject %>%
  select(EnrollmentID, ProjectName, ScoreAdjusted) %>%
  group_by(ProjectName) %>%
  summarise(AverageScore = round(mean(ScoreAdjusted), 1),
            EnrollmentCount = n())

CountyHousedAverageScores <- SPDATsByProject %>%
  group_by(CountyServed) %>%
  summarise(HousedAverageScore = round(mean(ScoreAdjusted), 1),
            HHsHousedInCounty = n())

gt(CountyHousedAverageScores)

```

## VI-SPDAT Scores Saved to Non-Heads-of-Households

If you have clients here, you should either verify the scores saved here are valid or the correct client is marked as the Head of Household.

```{r non_hohs_scores}
SPDATsOnNonHoHs <- left_join(Entries, Scores, by = "PersonalID") %>%
  filter(RelationshipToHoH != 1 & !is.na(Score)) %>%
  select(ProjectName, PersonalID, EntryDate, Score) %>%
  arrange(ProjectName)
gt(SPDATsOnNonHoHs)
```

## Map of Counties 

Each county is represented based on whether they are serving higher than the average SPDAT score in their county.

```{r}
library(maps)
Ohio <- map('county', 'ohio', fill = TRUE, col = palette("default"))

```



## Compare Average Scores Literally Homeless to Housed in a County

The following compares the Average VI-SPDAT Score of those who have entered a project that would mean they are literally homeless within a County to the Average VI-SPDAT Score of those who have an Entry into a Permanent Housing project type (RRH or PSH).

```{r compare}
Regions <- read_csv("public_data/Regions.csv") %>%
  mutate(RegionName = paste0("Homeless Planning Region ", Region))

Compare <- 
  full_join(CountyAverageScores, 
            CountyHousedAverageScores, 
            by = "CountyServed") %>%
  arrange(CountyServed) %>%
  left_join(., Regions, by = c("CountyServed" = "County")) %>%
  group_by(RegionName)

ggplot(Compare %>% filter(Region == 5), 
       aes(x = CountyServed, y = AverageScore)) + 
  geom_point(aes(x = CountyServed, y = AverageScore), size = 10, shape = 95) +
  theme(axis.text.x = element_text(size = 10)) +
  geom_point(aes(x = CountyServed, y = HousedAverageScore), 
             size = 4, 
             shape = 17) +
  xlab(Compare$RegionName)
```

