---
title: "How I Decided Which Providers to Use"
subtitle: "Coalition on Homelessness and Housing in Ohio"
author: "Genelle Denzin"
date: "9/9/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(readxl)
library(knitr)
```

## Data Sources

I'm beginning with the HUD CSV Export. I will only need the following files: Enrollment, Exit, Organization, and Project. The following code imports the data and assigns each dataset to a variable.

```{r getting_csv_data, message=FALSE, warning=FALSE}

Enrollment <- read_csv("data/Enrollment.csv")

Exit <- read_csv("data/Exit.csv")

Organization <- read_csv("data/Organization.csv")

Project <- read_csv("data/Project.csv")

```

## Further Adjustments to the HUD CSV Export 

Due to anamolies with ServicePoint's HUD CSV Export and data elements not required by HUD, but that are necessary for this analysis, I had to modify the contents of the imported tables from the HUD CSV Export in a few ways.

1.  Over the past year or so I've found numerous errors in ServicePoint's HUD CSV Export. The one that affects this analysis is that our Provider addresses are not being populated in the file. The columns are there, but the values are all null. The logic writes the Provider addresses from our custom reporting tool in ServicePoint into the file. 

2. In our vendor's custom reporting, the provider IDs of the Organizations come out tacked onto the end of the Organization names, so I have some logic there to remove that. 

3. Since Project County is not a HUD-required data element, the logic below adds the County in which the Project is located to the address data.

```{r csv_adjustments}

provider_data <- read_xlsx("data/RMisc.xlsx",
                            sheet = 4,
                            range = cell_cols("A:G")) %>%
  mutate(OrganizationName = str_remove(OrganizationName, "\\(.*\\)"))

Project <- Project %>%
  select(-ProjectName,-ProjectCommonName) %>%
  left_join(., provider_data, by = "ProjectID")

rm(provider_data)


```

4.  Since "County in which the Client is Served" is not a HUD-required data element, the following logic adds that data to Enrollment.csv. It is collected at Entry.

```{r}
counties <- read_xlsx("data/RMisc.xlsx",
                      sheet = 1,
                      range = cell_cols("B:E"),
                      col_types = c("numeric", "numeric", "text", "text"))

Enrollment <- left_join(Enrollment, counties, by = "EnrollmentID") 

rm(counties)
```

At this point, we have the data needed to complete the analysis.

## Finding the YHDP Projects to Include

### Prep

Here I am just setting up by assigning a Report Start and Report End and I copied a couple of functions I created for determining which projects were operating and which clients were served during the reporting period for use in this analysis. Also created an object to make it easier to call the Counties in question into later logic.

```{r set_up_objects}
ReportStart <- "10012018"
ReportEnd <- "09302019"

operating_between <- function(table, start, end) {
  operating <-  if_else(
    is.na(table$OperatingStartDate) |
      ymd(table$OperatingStartDate) > mdy(end) |
      (!is.na(table$OperatingEndDate) &
         ymd(table$OperatingEndDate) < mdy(start)),
    FALSE,
    TRUE
  )
  operating
}

served_between <- function(table, start, end){
  served <- ymd(table$EntryDate) <= mdy(end) &
    (is.na(table$ExitDate) | ymd(table$ExitDate) > mdy(start))
  served
}

YHDP_Counties <- c("Athens", "Jackson", "Vinton", "Gallia", "Meigs")
```

Just getting a slimmed down version of the Project.csv and renaming the County field so it's not confused with the CountyServed field, related to Enrollments.

```{r figuring}
small_projects <- Project %>%
  select(
    ProjectID,
    ProjectName,
    OrganizationName,    
    OrganizationID,
    OperatingStartDate,
    OperatingEndDate,
    ProjectType,
    ProjectCounty,
  )
```

### Which Projects are Located Within the Region?

We don't have any PH - Housing w/ Services, so I never include it in anything, but I would imagine other CoCs might want to include this Project Type. 

In deciding whether to include Services Only, my instinct was to exclude them since they weren't required for the general LSA, but after some discussion with Erica, I learned that the Services Only project in that region is funded by YHDP so we figure that project should actually be included.

I'm excluding Outreach because there are no PATH Street Outreach projects in the region, but I am hoping it will start to be included in HUD reporting so that we are not always missing our unsheltered clients. 

I'm excluding Prevention because those are not included in the LSA.

I decided not to filter out non-HMIS-participating projects, as we include those in the LSA.

I'm also limiting my results to those projects that were operational during the reporting period. (Using the Operating Start and End Dates.)

Finally, I want to filter down to the definite set of projects based on ProviderCounty, which comes from their address record.

```{r strict}
strictly_in_the_region <- small_projects %>%
  filter(ProjectType %in% c(1:3, 6, 8:9, 13) &
           operating_between(., ReportStart, ReportEnd) &
           ProjectCounty %in% YHDP_Counties) %>%
    select(ProjectID,
         ProjectName,
         OrganizationName,
         ProjectType,
         ProjectCounty)
```

### Which Projects Served Households in the Region?

Now that we have a core set of projects that will be included, I need to also gather the projects in the BoS that served households inside the YHDP counties in the reporting period.

I am filtering on similar things I filtered on above.

```{r served_in_region}
projects_serving_in_region <-
  Enrollment[c("PersonalID",
               "EnrollmentID",
               "EntryDate",
               "ProjectID",
               "CountyServed")] %>%
  left_join(Exit[c("PersonalID", 
                   "EnrollmentID",                     
                   "ExitDate")], by = c("PersonalID", "EnrollmentID")) %>%
  left_join(small_projects, by = "ProjectID") %>%
  filter(ProjectType %in% c(1:3, 6, 8:9, 13) &
           served_between(., ReportStart, ReportEnd) &
           CountyServed %in% YHDP_Counties) %>%
  select(OrganizationName,
         ProjectID,
         ProjectName,
         ProjectType,
         ProjectCounty) %>%
  unique()
```

### Combining the Lists

There are clearly some projects that exist outside the YHDP region that serve households in the region. So in order to wind up with a final list of projects to select for the LSA reporting group, we would add those to the projects that actually exist within the boundaries of the YHDP region.

```{r final}
final_list <- strictly_in_the_region %>%
  full_join(
    projects_serving_in_region,
    by = c(
      "ProjectID",
      "ProjectName",
      "OrganizationName",
      "ProjectType",
      "ProjectCounty"
    )
  ) %>%
  select(
    "Organization" = OrganizationName,    
    "Project ID" = ProjectID,
    "Project" = ProjectName,
    "Project Type" = ProjectType,
    "Project County" = ProjectCounty
  ) %>%
  arrange(Organization)

kable(final_list, format = 'markdown')
```

### Applying Local Knowledge

A thing I am noticing is that there is a project (1097) which is located outside our CoC who serves households in this region (and other BoS counties) that is pulling in. They enter BoS-based households into our HMIS and KY-based households into theirs. I'm leaving it in. This project was an issue for the 2017-2018 LSA and I wasn't able to get resolution from anyone at HUD about it.

Another issue I'm seeing is there are three site-based projects whose data indicates that they all served households in the YHDP region, but their sites are outside the region. So in these cases, the County Served data cannot be correct. I am going to manually remove those projects, with the suspicion that maybe in the projects_serving_in_region logic, it should have excluded projects where the Housing Type is not tenant-based.

## Final Results

Manually removing the three site-based projects discussed above.

```{r}
adjusted_final <- final_list %>%
  filter(!`Project ID` %in% c(25, 1202, 776))

kable(adjusted_final, format = 'markdown')
```

Thanks to Amanda Wilson and Erica Mulryan for helping with the Applying Local Knowledge part and reviewing this!

