---
title: "COVID-19 in Ohio BoS CoC Populations"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(knitr)
library(here)
library(tmap)    #for mapping
library(raster)  #for mapping

# FUNCTIONS
replace_yes_no <- function(column_name){
    if_else(column_name == "No" | is.na(column_name), 0, 1)
}

source(here("00_functions.R"))

# Specifying directories

if(exists("dataset") == FALSE) {
  dataset <- "live"
} else {
  dataset <- dataset
}

directory <- case_when(dataset == "live" ~ "data",
                       dataset == "sample" ~ "sampledata",
                       dataset == "yo" ~ "youngstowndata")

Enrollment <-
  read_csv(here(paste0(directory, "/Enrollment.csv")),
           col_types =
             "nnnDcnnnlnDnnnDDDnnnncccnnDnnnncnnnnnnnnnnnnnnnnnnnnnnnnnnnnnnnnnnTTnTn")

Exit <-
  read_csv(here(paste0(directory, "/Exit.csv")),
           col_types = "nnnDncnnnnnnnnnnnnnnnnnnnnnnnnnDnnnnnnTTnTn")

small_exit <- Exit %>% dplyr::select(EnrollmentID, 
                              ExitDate, 
                              Destination, 
                              OtherDestination)

Enrollment <- left_join(Enrollment, small_exit, by = "EnrollmentID") %>%
  mutate(ExitAdjust = if_else(is.na(ExitDate), today(), ExitDate))

rm(small_exit)

if(file.exists(here(paste0(directory, "/covid19.zip")))) {
  unzip(zipfile = here(paste0("./", directory, "/covid19.zip")),
        exdir = here(paste0("./", directory)))
  
  file.rename(here(paste0(
    directory, "/", list.files(here(paste0("./", directory)),
                               pattern = "(report_)")
  )),
  here(paste0(directory, "/covid19.csv")))
  
  file.remove(here(paste0(directory, "/covid19.zip")))
}

get_res_prior <- Enrollment %>%
  dplyr::select(PersonalID, EntryDate, ExitDate, LivingSituation) %>%
  group_by(PersonalID) %>%
  arrange(desc(EntryDate)) %>%
  slice(1L)

covid19 <- read_csv(here(paste0(directory, "/covid19.csv")),
                    col_types = "ncccccccccccccccccccccccccccccc") %>%
  mutate(
    COVID19AssessmentDate = mdy(COVID19AssessmentDate),
    ContactWithConfirmedDate = mdy(ContactWithConfirmedDate),
    ContactWithUnderInvestigationDate = mdy(ContactWithUnderInvestigationDate),
    TestDate = mdy(TestDate),
    DateUnderInvestigation = mdy(DateUnderInvestigation)
  ) %>%
  filter(ymd(COVID19AssessmentDate) > ymd("20200401") &
           ymd(COVID19AssessmentDate) <= today()) %>%
  left_join(get_res_prior, by = "PersonalID") %>%
  mutate(LivingSituationDescr = living_situation(LivingSituation)) %>%
  as_tibble() %>%
  mutate(
    Symptom1Cough = replace_yes_no(Symptom1Cough),
    Symptom1BreathingDifficult = replace_yes_no(Symptom1BreathingDifficult),
    Symptom2SoreThroat = replace_yes_no(Symptom2SoreThroat),
    Symptom2Fever = replace_yes_no(Symptom2Fever),
    Symptom2Chills = replace_yes_no(Symptom2Chills),
    Symptom2Headache = replace_yes_no(Symptom2Headache),
    Symptom2MusclePain = replace_yes_no(Symptom2MusclePain),
    Symptom2LostTasteSmell = replace_yes_no(Symptom2LostTasteSmell),
    Symptom2Congestion = replace_yes_no(Symptom2Congestion),
    Symptom2Nausea = replace_yes_no(Symptom2Nausea),
    Symptom2Diarrhea = replace_yes_no(Symptom2Diarrhea),
    Symptom2Weak = replace_yes_no(Symptom2Weak),
    HealthRiskChronicIllness = replace_yes_no(HealthRiskChronicIllness),
    HealthRiskHistoryOfRespiratoryIllness = replace_yes_no(HealthRiskHistoryOfRespiratoryIllness),
    HealthRiskOver65 = replace_yes_no(HealthRiskOver65),
    HealthRiskKidneyDisease = replace_yes_no(HealthRiskKidneyDisease),
    HealthRiskImmunocompromised = replace_yes_no(HealthRiskImmunocompromised),
    HealthRiskSmoke = replace_yes_no(HealthRiskSmoke),
    ContactWithConfirmedCOVID19Patient = replace_yes_no(ContactWithConfirmedCOVID19Patient),
    ContactWithUnderCOVID19Investigation = replace_yes_no(ContactWithUnderCOVID19Investigation),
    Tested = replace_yes_no(Tested),
    UnderInvestigation = replace_yes_no(UnderInvestigation)
  )

# living_situation() is a custom function I wrote that comes in with the COHHIO_HMIS data. (see line 13) It just translates the number assigned to the various responses to the descriptions acc. to the HUD Data Standards.

rm(list = ls()[!(ls() %in% c('covid19',
                             'entered_between',
                             'exited_between',
                             'served_between',
                             'stayed_between'))])
```

## What Data We're Collecting

The Ohio Balance of State CoC is collecting COVID-19 data based on the CDC's guidelines, listed [here](https://www.cdc.gov/coronavirus/2019-ncov/symptoms-testing/symptoms.html). We began collecting this data in April 2020.

Please send inquiries to hmis@cohhio.org.

## COVID-19 Status of Clients Over Time

Since April 1, **`r nrow(covid19)`** clients have been screened for COVID-19. "No Current Indications" means the client does not have symptoms that would indicate they may have COVID-19, they haven't reported any recent contact with an infected person, and they have not tested positive. "May Have COVID-19" means the client fits what is listed in the [CDC's guidelines](https://www.cdc.gov/coronavirus/2019-ncov/symptoms-testing/symptoms.html) as either having symptoms or has had contact with a person who is confirmed to have or is under investigation for COVID-19. "Positive" means the client has self-reported that they were tested for COVID-19 by a medical professional within the last 14 days and the test was positive. This plot does not include data about risk factors.

```{r fig1, fig.asp=.62, fig.width=8, message=FALSE}
# since this is meant to show the state of clients over time and at the point of their most recent assessment, this logic is a little different from the logic for the Prioritization report in that we are throwing out symptom data based on the AssessmentDate, not based on its distance from "today".

assessments_over_time_plot <- covid19 %>%
  mutate(
    COVID19Status = case_when(
      Tested == 1 &
        TestResults == "Positive" &
        ymd(TestDate) > ymd(COVID19AssessmentDate) - days(14) &
        !is.na(TestDate) ~ "Positive",
      # testing positive in the 14 days prior to assessment is the only way to
      # land in this bucket
      (
        ContactWithConfirmedCOVID19Patient == 1 &
          (
            ymd(ContactWithConfirmedDate) >
              ymd(COVID19AssessmentDate) - days(14) |
              is.na(ContactWithConfirmedDate)
          )
        # compares contact date to date of the assessment
      ) |
        (
          ContactWithUnderCOVID19Investigation == 1 &
            (
              ymd(ContactWithUnderInvestigationDate) >
                ymd(COVID19AssessmentDate) - days(14) |
                is.na(ContactWithUnderInvestigationDate)
            )
        ) |
        (
          Symptom1BreathingDifficult +
            Symptom1Cough +
            Symptom2Chills +
            Symptom2SoreThroat +
            Symptom2Fever +
            Symptom2Headache +
            Symptom2LostTasteSmell +
            Symptom2MusclePain +
            Symptom2Congestion +
            Symptom2Nausea +
            Symptom2Diarrhea +
            Symptom2Weak
        ) > 0
      |
        (
          UnderInvestigation == 1 &
            ymd(DateUnderInvestigation) > ymd(COVID19AssessmentDate) - days(14)
        ) ~
        "May Have COVID-19",
      # being Under Investigation (past 14 days), any Symptom, or any Contact
      # in the 14 days prior to the assessment date will land you here ^
      TRUE ~ "No Current Indications"
      # everyone else lands here ^
    ),
    COVID19Status = factor(
      COVID19Status,
      levels = c("No Current Indications",
                 "May Have COVID-19",
                 "Positive")
    )
  )

assessments_over_time_plot %>%
  dplyr::select(PersonalID, COVID19AssessmentDate, COVID19Status) %>%
  group_by(COVID19AssessmentDate, COVID19Status) %>%
  summarise(Clients = n()) %>%
  ggplot(aes(x = COVID19AssessmentDate, y = Clients,
             fill = COVID19Status)) +
  geom_bar(stat = "identity", 
           position = position_stack(reverse = TRUE)) +  
  scale_fill_brewer(palette = "Dark2") +
  theme_classic() +
  ggtitle("Clients COVID-19 Assessments in the Ohio Balance of State CoC") +
  labs(x = "Date of Assessment", y = "Clients Assessed") +
  theme(legend.title=element_blank(),
        legend.position = "top",
        legend.key.height = unit(0.1, "cm"),
        legend.key.width = unit(0.3, "cm"))

```

\pagebreak

## Priority for Immediate Non-congregate Housing

```{r priority, message=FALSE}

priority <- covid19 %>%
  mutate(
    Priority = case_when(
      # if tested positive
      (
        Tested == 1 &
          TestResults == "Positive" &
          ymd(TestDate) > ymd(COVID19AssessmentDate) - days(14) &
          !is.na(TestDate)
      ) |
        # if under investigation
        (
          UnderInvestigation == 1 &
            ymd(DateUnderInvestigation) > ymd(COVID19AssessmentDate) - days(14)
        ) |
        # contact with COVID-19
        (
          ContactWithConfirmedCOVID19Patient == 1 &
            (
              ymd(ContactWithConfirmedDate) >
                ymd(COVID19AssessmentDate) - days(14) |
                is.na(ContactWithConfirmedDate)
            )
          # compares contact date to the assessment date too since we want to
          # see severity at the time of assessment
        ) |
        (
          ContactWithUnderCOVID19Investigation == 1 &
            (
              ymd(ContactWithUnderInvestigationDate) >
                ymd(COVID19AssessmentDate) - days(14) |
                is.na(ContactWithUnderInvestigationDate)
            )
        ) |
        # if the client came from jail or nursing home
        (
          LivingSituation %in% c(7, 25) &
            EntryDate > ymd(COVID19AssessmentDate) - days(14) &
            EntryDate <= ymd(COVID19AssessmentDate)
        ) |
        # if the client has any symptoms at all
        (
          Symptom1BreathingDifficult +
            Symptom1Cough +
            Symptom2Chills +
            Symptom2SoreThroat +
            Symptom2Fever +
            Symptom2Headache +
            Symptom2LostTasteSmell +
            Symptom2MusclePain +
            Symptom2Congestion +
            Symptom2Nausea +
            Symptom2Diarrhea +
            Symptom2Weak
        ) > 0 ~ "Needs Isolation/Quarantine",
        # if the client has any risks at all
        (
          HealthRiskHistoryOfRespiratoryIllness +
            HealthRiskChronicIllness +
            HealthRiskOver65 +
            HealthRiskKidneyDisease +
            HealthRiskImmunocompromised +
            HealthRiskSmoke > 0
        )  ~ "Has Health Risk(s)",
      TRUE ~ "No Known Risks or Exposure"
      # everyone else lands here ^
      # in the report, there will be a third level: "Not Assessed Recently"
    ),
    Priority = factor(Priority, levels = c("Needs Isolation/Quarantine", 
                                           "Has Health Risk(s)", 
                                           "No Known Risks or Exposure"))
  )

```

Our Prioritization report for Coordinated Entry will use this metric to determine whether a client should be prioritized for immediate non-congregate housing based on their risks, symptoms, test results, and contact with others. This report is used to help communities with moving households into permanent housing. 

- The **Needs Isolation/Quarantine** category indicates the household should be moved out of any congregate space immediately for quarantine due to a suspected or confirmed infection or likely exposure. Likely exposure could include having come into the system directly from prison or a nursing home.
- The **Has Health Risk(s)** category indicates the household needs non-congregate sheltering as the person is at high risk of poor outcomes if infected with COVID-19 due to health risks that match with CDC's guidelines. 
- Although everyone is at risk of contracting this virus when exposed, the **No Known Risks or Exposure** category means the household has no known risk factors, and no suspected or known exposure.

```{r fig2, fig.asp=.62, fig.width=8, message=FALSE}

priority_plot <- priority %>%
  dplyr::select(PersonalID, COVID19AssessmentDate, Priority) %>%
  group_by(COVID19AssessmentDate, Priority) %>%
  summarise(Clients = n()) %>%
  arrange(Priority)

priority_plot %>%
  ggplot(aes(x = COVID19AssessmentDate, y = Clients,
             fill = Priority, label = Clients)) +
  scale_fill_brewer(palette = "GnBu", direction = -1) +
  geom_bar(stat = "identity") +
  theme_classic() +
  ggtitle("Priority for Immediate Housing Based on Symptoms and Risk Factors") +
  theme(legend.title=element_blank(),
        legend.position = "top",
        legend.key.height = unit(0.1, "cm"),
        legend.key.width = unit(0.3, "cm"))+
  labs(x = "Date of Assessment", y = "Clients Assessed")

```

\pagebreak

## Geography

### Where are clients being assessed?
This map shows the total number of clients assessed with the COVID Screening Tool  and reported in HMIS by county.

```{r geo, dpi=300, echo=FALSE, message=FALSE}

covid19_county_summary <- covid19 %>%
  group_by(CountyServed) %>%
  summarise(ClientCount = n()) %>%
  dplyr::select("NAME_2" = CountyServed, "Clients Screened" = ClientCount)

# Get polygon data
us <- getData('GADM', country='USA', level=2)
ohio_shp <- subset(us, NAME_1 == "Ohio")
counties_shp <- subset(ohio_shp, NAME_2 != "Lake Erie")


counties_covid_shp <- merge(counties_shp, covid19_county_summary,
                            by = "NAME_2")

tmap_mode("plot")
qtm(counties_covid_shp,
    fill = "Clients Screened",
    fill.palette = "Blues",
    frame = FALSE,
    text = "NAME_2",
    text.size = .35,
    legend.outside = TRUE
    )

```
