---
title: "Orgs Coming to BitFocus"
author: "Genelle Denzin"
date: "2/27/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, include=FALSE, message=FALSE}

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(lubridate)
library(here)
library(readxl)
library(gt)

```

## Migration to BitFocus

This document aims to identify all Organizations in our system and separate them into three categories: Not Migrating, Migrating Active, and Migrating Inactive. BitFocus has agreed to waive fees for Organizations that have no Projects that are HMIS participating, so we need to definitively name those organizations for the contract.

```{r gather}

art_data <-
  read_xlsx(here("random_data/OrganizationsBitFocus.xlsx"),
            sheet = 4) %>%
  filter(!is.na(ProjectID) & ProjectID != 1) %>% # dropping deleted providers and Provider ID 1
  mutate(
    MinEntry = as.Date(MinEntry, origin = "1899-12-30"),
    MaxExit = as.Date(MaxExit, origin = "1899-12-30"),
    OpenEEs = replace_na(OpenEEs, 0),
    operating = if_else(operating == "Yes", 1, 0),
    participating = if_else(participating == "Yes", 1, 0),
    ce_or_vash = if_else(
      str_detect(Project, "VASH") | ptc == "Coordinated Entry (HUD)",
      1,
      0
    ),
    obsolete_ptc = if_else(
      ptc %in% c(
        "Family Homelessness Prevention Pilot (FHPP)",
        "RETIRED (HUD)",
        "Homelessness Prevention",
        "Housing Stability Program (HSP)",
        "Direct Housing"
      ),
      1,
      0
    )
  )

raw_data_projects <- nrow(art_data)

obsolete_count <- nrow(art_data %>%
                         filter(obsolete_ptc == 1))

ce_or_vash_count <- nrow(art_data %>%
                           filter(ce_or_vash == 1))

participating_count <- nrow(art_data %>%
                           filter(participating == 1))

operating_count <- nrow(art_data %>%
                           filter(operating == 1))

operating_and_participating_count <- nrow(art_data %>%
                                            filter(operating == 1 &
                                                     participating == 1))

operating_not_participating_count <- nrow(art_data %>%
                                            filter(operating == 1 &
                                                     participating == 0))

participating_not_operating <- nrow(art_data %>%
                                         filter(operating == 0 &
                                                   participating == 1))

open_ees_count <- nrow(art_data %>%
                         filter(OpenEEs > 0))

```


## Projects and Organizations {.tabset .tabset-pills}

### Raw Data

We are starting with a `r raw_data_projects`-row dataset that includes all providers in ServicePoint minus deleted providers. 

Some identifiers in the data from ServicePoint to help categorize the organizations go like this:

  - `r ce_or_vash_count` APs or VASH providers - will categorize as Migrating Active
  - `r obsolete_count` Providers with obsolete Project Types (Direct Housing, the non-HUD Homelessness Prevention, etc.) - will categorize as Not Migrating
  - `r open_ees_count` Providers with open Entry Exits - with some exceptions, will categorize as Migrating Active
  - `r participating_count` HMIS Participating - will categorize as Migrating Active
  - `r operating_count` Operating - will categorize as Migrating Active or Inactive, depending on other factors
  
```{r}
gt(art_data) %>%
  tab_header(
    title = "ALL Providers Except Deleted & Project ID 1",
    subtitle = paste(raw_data_projects, "records")
  ) %>%
  tab_source_note(
    source_note = "Source: ART report in Public > SysAdmin > Organizations for BitFocus"
  )  %>%
  tab_options(
    table.width = pct(100),
    table.font.size = pct(70)
  )
```

### Grouped by Organization

Now we want to look within each Organization to determine if it needs to be migrated, and if so, whether it should be considered Active or Inactive. 

Here we can see across all projects associated with a given Organization: the earliest Entry Date, the most recent Exit Date, whether there are any open Entry Exits, whether any projects within the organization are marked as Operating or Participating, and whether any projects are Coordinated Entry or VASH projects. We are going to use this data in further calculations. No records have been lost here, only grouped so that each Organization gets one row.

```{r grouping, warning=FALSE}

org_level <- art_data %>%
  group_by(Org) %>%
  summarise(minEntry = min(ymd(MinEntry), na.rm = TRUE),
            maxExit = max(ymd(MaxExit), na.rm = TRUE),
            openEEs = sum(OpenEEs),
            maxOperating = max(operating),
            maxParticipating = max(participating),
            CEorVASH = max(ce_or_vash))

orgs <- nrow(org_level)

```

At this point we have `r orgs` records in the dataset, representing `r raw_data_projects` projects.

```{r}

gt(org_level) %>%
  tab_header(
    title = "ALL Organizations",
    subtitle = paste(orgs, "records")
  ) %>%
  tab_options(
    table.width = pct(100),
    table.font.size = pct(70)
  )

```


### Not Migrating

These Organizations are not migrating. I'm defining this by looking at all organizations where there are no projects associated with them that are HMIS-participating, VASH, CE, there are no open Entry Exits, and the most recent Exit Date is before 5/1/2014 or there was no data entered ever.

```{r notmigratingcount}

not_migrating <- org_level %>%
  mutate(nodata = str_detect(minEntry, "/")) %>%
  filter((ymd(maxExit) <= ymd("20140501") | is.na(nodata)) &
            openEEs == 0 &
           maxParticipating == 0 &
           maxOperating == 0 &
           CEorVASH == 0) %>%
  unique() 

not_migrating_count <- nrow(not_migrating)

```

We have at least `r not_migrating_count` organizations we do not need to migrate to Clarity.

```{r notmigrating}
gt(not_migrating %>% select(-nodata)) %>%
  tab_header(
    title = "Not Migrating",
    subtitle = paste(not_migrating_count, "records")
  ) %>%
  tab_options(
    table.width = pct(100),
    table.font.size = pct(70)
  )

```

### Migrating Inactive
```{r migratinginactive}
possibly_migrating <- org_level %>%
  anti_join(not_migrating, by = "Org")

migrating_inactive <- possibly_migrating %>%
  mutate(nodata = str_detect(minEntry, "/")) %>%
  filter(maxParticipating == 0 &
           CEorVASH == 0 &
           openEEs == 0)

possibly_migrating_count <- nrow(possibly_migrating)
migrating_inactive_count <- nrow(migrating_inactive)

```

Next we are examining the `r possibly_migrating_count` remaining Organizations to determine, since they are migrating, whether we should consider them Active or Inactive. After we removed the records above, now we are left with those who either have EEs in the past 7 years or have no data but maybe a VASH, CE, an HMIS-participating, or operating project.

Out of `r possibly_migrating_count` organizations that are possibly migrating to Clarity, `r migrating_inactive_count` are being counted here as inactive, or "shell" organizations.

```{r}
gt(migrating_inactive %>% select(-nodata)) %>%
  tab_header(
    title = "Migrating Inactive",
    subtitle = paste(migrating_inactive_count, "records out of", possibly_migrating_count,
                     "migrating")
  ) %>%
  tab_options(
    table.width = pct(100),
    table.font.size = pct(70)
  )

```


### Migrating Active

Theoretically, what should be left in the list are all the organizations we need to migrate and consider to be active for the purposes of the contract. This should include organizations with active data, Access Point projects, and VASH projects.

```{r migratingactive}
migrating_active <- possibly_migrating %>%
  anti_join(migrating_inactive, by = "Org")

migrating_active_count <- nrow(migrating_active)
```

Out of `r possibly_migrating_count` organizations that are possibly migrating to Clarity, `r migrating_active_count` are being counted here as active organizations.

```{r}
gt(migrating_active) %>%
  tab_header(
    title = "Migrating Active",
    subtitle = paste(migrating_active_count, "records out of", possibly_migrating_count,
                     "migrating")
  ) %>%
  tab_options(
    table.width = pct(100),
    table.font.size = pct(70)
  )

```

### Maybe Add to Migrating Inactive

```{r}

level2_providers <- art_data %>%
  filter(Org == "Coalition on Homelessness and Housing in Ohio(1)" &
           ptc == "Administrating Agency" &
           str_starts(Project, "zz", negate = TRUE)) %>%
  select("Org" = Project)

referral_orgs <-
  anti_join(level2_providers,
            art_data,
            by = "Org") %>%
  arrange(Org)

referral_orgs_count <- nrow(referral_orgs)

```

There are `r referral_orgs_count` projects that have no child-providers which could be considered to be Organizations, even if they do not show as Organizations in ServicePoint technically. This lists those projects that are not zz'd.

If we are wanting to add these into the Migrating Inactive category, then these would add to the `r migrating_inactive_count` other organizations we are already counting as Migrating Inactive.

```{r}
gt(referral_orgs)  %>%
  tab_header(
    title = "Maybe Add to Migrating Inactive",
    subtitle = paste(referral_orgs_count, "records")
  ) %>%
  tab_options(
    table.width = pct(100),
    table.font.size = pct(70)
  )
```



## {-}



